---
title: "Tiny practices with gt"
author: "Allison Horst"
date: "January 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I made this because I needed to do some really simple examples. Here are 10 different gt tables that worked, using short examples with data existing in the *datasets* package or that I create here. 

### Attach packages
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(gt)
```

###Example 0. It just works... 

Make a table of the first 5 observations in the 'rock' dataset.

```{r}

rock %>% # Get 'rock' data
  head(5) %>% # First 5 lines only
  gt() # Make a table, it just works. 

```

###Example 1. The first try at customization

Biological oxygen demand table (BOD)

Use the built-in dataset BOD to create a simple customized table. The BOD dataset (see ?BOD for documentation) has two columns: **Time** and **demand** (both numeric).

```{r}

BOD %>% # Get the data...
gt() %>% # use 'gt' to make an awesome table...
  tab_header( 
    title = "BOD Table Woooooo!", # ...with this title
    subtitle = "Hooray gt!") %>% # and this subtitle
  fmt_number( # A column (numeric data)
    columns = vars(Time), # What column variable? BOD$Time
    decimals = 2 # With two decimal places
    ) %>% 
  fmt_number( # Another column (also numeric data)
    columns = vars(demand), # What column variable? BOD$demand
    decimals = 1 # I want this column to have one decimal place
  ) %>% 
  cols_label(Time = "Time (hours)", demand = "Demand (mg/L)") # Update labels

``` 

###Example 2. The one where I added colors and footnotes 

ToothGrowth (VitC effect on tooth growth in guinea pigs)

See ?ToothGrowth for dataset documentation. The ToothGrowth dataset has three variables: **len** (tooth length), **supp** (supplement ID), **dose** (dosage). I use group_by + summarize here to find mean tooth length grouped by supplement and dosage first, then create a table of the means. 

```{r}

# Some wrangling (grouped means by supplement + dosage):

tooth_length <- ToothGrowth %>% 
  group_by(supp, dose) %>% 
  summarize(
    mean_len = mean(len)
  ) %>% 
  as_tibble() 

# A gt table: 
tooth_length %>% # Take tooth_length
  gt() %>% # Make a gt table with it
  tab_header(
    title = "A title just like that", # Add a title
    subtitle = "(with something below it!)" # And a subtitle
  ) %>%
  fmt_passthrough( # Not sure about this but it works...
    columns = vars(supp) # First column: supp (character)
  ) %>% 
  fmt_number(
    columns = vars(mean_len), # Second column: mean_len (numeric)
    decimals = 2 # With 4 decimal places
  ) %>% 
  fmt_number(
    columns = vars(dose), # Third column: dose (numeric)
    decimals = 2 # With 2 decimal places
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(supp), # ...for supp column!
    colors = scales::col_factor( # <- bc it's a factor
      palette = c(
        "green","cyan"), # Two factor levels, two colors
      domain = c("OJ","VC")# Levels
  )
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(dose), # ...for dose column 
    colors = scales::col_numeric( # <- bc it's numeric
      palette = c(
        "yellow","orange"), # A color scheme (gradient)
      domain = c(0.5,2) # Column scale endpoints
  )
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(mean_len), # ...for mean_len column
    colors = scales::col_numeric(
      palette = c(
        "red", "purple"), # Overboard colors! 
      domain = c(7,27) # Column scale endpoints
  )
  ) %>% 
  cols_label(supp = "Supplement", dose = "Dosage (mg/d)", mean_len = "Mean Tooth Length") %>% # Make the column headers
  tab_footnote(
    footnote = "Baby footnote test", # This is the footnote text
    locations = cells_column_labels(
      columns = vars(supp) # Associated with column 'supp'
      )
    ) %>% 
    tab_footnote(
    footnote = "A second footnote", # Another line of footnote text
    locations = cells_column_labels( 
      columns = vars(dose) # Associated with column 'dose'
      )
    )

```

###Example 3. Updating NAs, hiding columns

Using first 10 observations in the 'airquality' dataset. Existing variables: **Ozone**, **Solar.R**, **Wind**, **Temp**, **Month**, **Day** 

```{r}

airquality %>% # Take the 'airquality' dataset...
  head(10) %>% # Then just grab the first 10 observations...
  gt() %>% # Then make a gt table
  tab_header(
    title = "Air Quality dataset", # Add a title
    subtitle = "...also with a subtitle" # And a subtitle
  ) %>%
  fmt_missing( # Then reassign the 'NA' values to something new
    columns = vars(Ozone), # For variable 'Ozone', make "NA"s
    missing_text = "Teddy" # ...to "Teddy" instead
  ) %>% 
  fmt_missing( # Reassign NAs for another column
    columns = vars(Solar.R), # For variable 'Solar.R', make "NA"s
    missing_text = "Dog" # ...to "Dog" instead
  ) %>% 
  cols_hide( # Then hide the 'Wind' and 'Month' variables
    columns = vars(Wind, Month)
  )

```

###Example 4. Scientific notation (...and more colors, can't help it...) 

I'll just make some data here (tibble 'fake'). PS this is my first time using tribble and it is SHOCKINGLY simple. What have I even been doing with my life... 

```{r}

fake <- tribble(
  ~Species, ~Height, ~Mass,
  "Blarg", 20500, 1980,
  "Gorple", 17000, 775,
  "Roglob", 11820, 20,
  "Fwerpy", 8005, 352
) %>% 
  arrange(-Mass)

```

Then make a little table of 'fake' with values in scientific notation:

```{r}

fake %>% 
  gt() %>% 
  tab_header(
    title = "Fake species data", # Add a title
    subtitle = "another useful subtitle" # And a subtitle
  ) %>%
  fmt_scientific( # Reformat to scientific notation...
    columns = vars(Height), # The values for the 'Height' variable
    decimals = 3 # Keeping 3 decimal places
  ) %>% 
  fmt_scientific( # Reformat to scientific notation...
    columns = vars(Mass), # The values for the 'Mass' variable
    decimals = 1 # Keeping 1 decimal place
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(Mass), # ...for Mass column 
    colors = scales::col_numeric( # <- bc it's numeric
      palette = c(
        "yellow", "red", "purple"), # A color scheme (gradient)
      domain = c(1e1, 2e3) # Column scale endpoints
  )
  )
  

```

###Example 5. Date formatting + column spanning labels

First, I'll make a little tibble with dates (again using tribble, my new favorite thing):

```{r}

date_trial <- tribble(
  ~Date, ~Temp, ~Rain, ~Flavor, 
  "1980-01-04", 75, 0.4, "Grape",
  "1975-05-28", 80, 0.2, "Strawberry",
  "1993-09-02", 94, 0.1, "Chocolate"
)

```

Then into a nice table: 

```{r}

date_trial %>% 
  gt() %>% 
  fmt_date(
    columns = vars(Date),
    date_style = 5 # Try different numbers here to reformat! Cool!
  ) %>% 
  tab_spanner(
    label = "Weather",
    columns = vars(Temp, Rain)
  )

```

