---
title: "Tiny practice with gt"
author: "Allison Horst"
date: "January 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I made this because I needed to do some really simple examples. Here are 10 different gt tables that worked, using short examples with data existing in the *datasets* package. 

### Attach packages
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(gt)
```

###Ground Zero. It just works. 

Make a table of the first 5 observations in the 'rock' dataset.

```{r}

rock %>% # Get 'rock' data
  head(5) %>% # First 5 lines only
  gt() # Make a table, it just works. 

```

###Example 1. The first try at customization

Biological oxygen demand table (BOD)

Use the built-in dataset BOD to create a simple customized table. The BOD dataset (see ?BOD for documentation) has two columns: **Time** and **demand** (both numeric).

```{r}

BOD %>% # Get the data...
gt() %>% # use 'gt' to make an awesome table...
  tab_header( 
    title = "BOD Table Woooooo!", # ...with this title
    subtitle = "Hooray gt!") %>% # and this subtitle
  fmt_number( # Column 1 (numeric data)
    columns = vars(Time), # What column 1 variable? BOD$Time
    decimals = 2 # With two decimal places
    ) %>% 
  fmt_number( # Column 2 (also numeric data)
    columns = vars(demand), # What column 2 variable? BOD$demand
    decimals = 1 # I want this column to have one decimal place
  ) %>% 
  cols_label(Time = "Time (hours)", demand = "Demand (mg/L)") # Update labels

``` 

###Example 2. The one where I added colors and footnotes 

ToothGrowth (VitC effect on tooth growth in guinea pigs)

See ?ToothGrowth for dataset documentation. The ToothGrowth dataset has three variables: **len** (tooth length), **supp** (supplement ID), **dose** (dosage). I use group_by + summarize here to find mean tooth length grouped by supplement and dosage first, then create a table of the means. 

```{r}

# Some wrangling (grouped means by supplement + dosage):

tooth_length <- ToothGrowth %>% 
  group_by(supp, dose) %>% 
  summarize(
    mean_len = mean(len)
  ) %>% 
  as_tibble() 

# A gt table: 
tooth_length %>% # Take tooth_length
  gt() %>% # Make a gt table with it
  tab_header(
    title = "A title just like that", # Add a title
    subtitle = "(with something below it!)" # And a subtitle
  ) %>%
  fmt_passthrough( # Not sure about this but it works...
    columns = vars(supp) # First column: supp (character)
  ) %>% 
  fmt_number(
    columns = vars(mean_len), # Second column: mean_len (numeric)
    decimals = 2 # With 4 decimal places
  ) %>% 
  fmt_number(
    columns = vars(dose), # Third column: dose (numeric)
    decimals = 2 # With 2 decimal places
  ) %>% 
  data_color(
    columns = vars(supp), # Cell colors for supp column
    colors = scales::col_factor( # <- bc it's a factor
      palette = c(
        "green","cyan"), # Two factor levels, two colors
      domain = c("OJ","VC")# Levels
  )
  ) %>% 
  data_color(
    columns = vars(dose), # Cell colors for dose column 
    colors = scales::col_numeric( # <- bc it's numeric
      palette = c(
        "yellow","orange"), # A color scheme (gradient)
      domain = c(0.5,2) # Column scale endpoints
  )
  ) %>% 
  data_color(
    columns = vars(mean_len), # Cell colors for mean_len column
    colors = scales::col_numeric(
      palette = c(
        "red", "purple"), # Overboard colors! 
      domain = c(7,27) # Column scale endpoints
  )
  ) %>% 
  cols_label(supp = "Supplement", dose = "Dosage (mg/d)", mean_len = "Mean Tooth Length") %>% # Make the column headers
  tab_footnote(
    footnote = "Baby footnote test", # This is the footnote text
    locations = cells_column_labels(
      columns = vars(supp) # Associated with column 'supp'
      )
    ) %>% 
    tab_footnote(
    footnote = "A second footnote", # Another line of footnote text
    locations = cells_column_labels( 
      columns = vars(dose) # Associated with column 'dose'
      )
    )

```

###Example 3. Trying updating NAs and hiding a column

Using first 10 observations in the 'airquality' dataset. Existing variables: **Ozone**, **Solar.R**, **Wind**, **Temp**, **Month**, **Day** 

```{r}

airquality %>% 
  head(10) %>% 
  gt()

```



